Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-01 cluster_paper/ms.tex
update ms

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
  cluster orbit.  In different projections, the line spans different
  time intervals: the longest time window is visible in the $(Z,X)$
  projection.  The large blue arrows denote the median cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By comparison,
  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the inertial
  frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
  direction.
  The sizes of the blue arrows are internally proportionate. 
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
     The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $Bp-Rp\approx$2.2).
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
members of the cluster show a clean sequence consistent with stars
with a fixed age and metallicity, and varying mass.  The halo members
mostly show a consistent main-sequence that has somewhat greater
scatter.  One possible explanation for the additional scatter could be
that the halo is more contaminated by field stars.  For instance,
$\approx5$ evolved halo stars are visible in the field-star portion of
the red giant branch (RGB).  Given the age of the system, these stars
must be field interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 300\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

@@

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  However below this mass, the data and models diverge
at colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Explanations invoking both starspots and
incomplete molecular line lists have been proposed \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal predictive
power.

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  The statistical uncertainties ($\approx
25$\,Myr) were comparable to the systematic uncertainties.

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.
This implies an absolute age slightly older than the Pleiades
\citep[cf.][]{mermilliod_comparative_1981}.

% FIXME: do you want to do the procedure below? Seems like overkill
% given the Cummings study.

% Our photometric age-dating procedure is as follows.  We selected all
% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
% outliers from the rest of the cluster.  Using an isochrone grid
% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
% spacing, we then numerically evaluated the $\chi^2$ fit for each
% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
% percentiles are used as summary statstics.
% 
% % NOTE: probably asymmetric.
% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
% yields $XXX\pm YYY$\,Myr.  Finally, applying the same procedure to
% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
% [nb. the procedure should be simple and general enough that you can
% apply it to the field sample, and it'll just give a broad
% distribution...]

@@

        \vspace{-0.1cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Subset
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right columns}) undergo
    successive degrees of automated cleaning, described in
    Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$,
    in which some NGC\,2516 members appear slightly elevated with respect to
    the Pleiades, suggesting an older age.
    {\it Bottom row}: \cn\ colored by binarity indicators.
    Sources with $\mathrm{RUWE}>1.2$ appear in red (astrometric
    binaries).
    Overplotted in orange are sources at least 0.3\,mag above an
    empirically fit cluster isochrone (photometric binaries).
		\label{fig:rot}
	}
\end{figure*}

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, following three different sets of cleaning criteria.  The
first and second sets of criteria were entirely automated, and yielded
subsets we denote $\mathcal{A}$ and $\mathcal{B}$.  For both subsets,
we considered light curves for which the peak Lomb-Scargle periodogram
period was below 15 days, the normalized Lomb-Scargle power exceeded
0.08, and for which no equal-brightness or greater companions were
within the aperture.  Beyond requiring that the target star be the
brightest star in the aperture, we also required that at most one
companion with brightness exceeding one tenth of the target star's
brightness be present in the aperture according to the Gaia DR2 source
catalog.  For set $\mathcal{A}$, this yielded \nautorotnumerator light
curves, out of \nautorotdenominator stars for which rotation periods
might plausibly have been detected.  For set $\mathcal{B}$, we 
additionally required that the Lomb-Scargle and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%.
This yielded \nautorotnumeratormatching light curves.

The comparison to the Pleiades and Praesepe is made using the rotation
periods and dereddened colors from Table~5 of
\citet{curtis_rup147_2020}, which used data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.
A few conclusions can be drawn by comparing the \cn\ rotation periods
with those of the Pleiades and Praesepe.  The first is that in both
stages of cleaning, the core and halo of \cn\ both show
gyrochronological ``slow sequences''.  These sequences appear to be
coeval with each other, and at least to first order appear to overlap
with the Pleiades sequence.

Second, subset $\mathcal{A}$ is more complete than subset
$\mathcal{B}$.  This completeness comes at the expense of greater
contamination.  The slow sequence has a factor of $\sim$twice as many
stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
the slow sequence, which could well be field contaminants, for the
most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
the slow sequence is also more pronounced in the cleaner subset,
though it seems to not be as empty as in the Pleiades.

Finally, comparing the slow sequences of the Pleaides and \cn, they
appear to overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  For stars that
are redder, the dispersion in rotation periods increases.  From
$1.2<(Bp-Rp)_0<1.7$, subset $\mathcal{B}$ of \cn\ has stars that
overlap with the Pleiades.  However, it also has about ten stars at
longer rotation periods---extending up to $\approx$11 days, rather
than the $\approx$8.5day limit for the Pleiades.  This is consistent
with \cn\ being slightly older than the Pleiades.
{\bf We explore this quantitatively by fitting interpolated
models}[?].  The resulting gyrochronology age we find for the core is
XXX.  For the halo, the claimed age from gyrochronology is YYY.
Applying the same procedure to the field star comparison sample (see
Appendix~\ref{app:compstar}), we get an age of ZZZ+/-AAA.

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
  cluster orbit.
  In different projections, the gray line spans different
  time intervals: the longest time window is visible in the $(Z,X)$
  projection.  The large blue arrows denote the median cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By comparison,
  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the inertial
  frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
  direction.
  The sizes of the blue arrows are internally proportionate. 
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
    are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
members of the cluster show a clean sequence consistent with stars
with a fixed age and metallicity, and varying mass.  The halo members
show a consistent sequence, though with greater scatter.  One possible
explanation for the additional scatter could be that the halo is more
contaminated by field stars.  For instance, $\approx5$ halo
stars are present in portions of the red giant branch
(RGB) that imply stellar ages inconsistent with that of the cluster.
These stars must be field interlopers.

A second possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 300\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

@@

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  Below this mass, the data and models diverge at
colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.\footnote{The 100 Myr isochrone does overlap well with the
lower main sequence, but it fails to fit the upper main sequence and
red giants.}  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
between the models and observations have include starspots and
incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal
predictive power.

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  Their statistical uncertainties ($\approx
25$\,Myr) were comparable to the systematic uncertainties.

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The RGB stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for a more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt an PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of their quoted statistical
and systematic components.  The relative age of ``slightly older than
the Pleiades'' is likely the more robust statement, consistent with
reports since \citet{mermilliod_comparative_1981}.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Subset
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.
    {\it Bottom row}: \cn\ colored by binarity indicators.  Sources
    with $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\%
    of stars).
    Overplotted in orange are sources at least 0.3\,mag above an
    empirically fit cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, following three different sets of cleaning criteria.  The
first and second sets of criteria were entirely automated, and yielded
subsets we denote $\mathcal{A}$ and $\mathcal{B}$.  For both subsets,
we considered light curves for which the peak Lomb-Scargle periodogram
period was below 15 days, the normalized Lomb-Scargle power exceeded
0.08, and for which no equal-brightness or greater companions were
within the aperture.  Beyond requiring that the target star be the
brightest star in the aperture, we also required that at most one
companion with brightness exceeding one tenth of the target star's
brightness be present in the aperture according to the Gaia DR2 source
catalog.  For set $\mathcal{A}$, this yielded \nautorotnumerator light
curves, out of \nautorotdenominator stars for which rotation periods
might plausibly have been detected.  For set $\mathcal{B}$, we 
additionally required that the Lomb-Scargle and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%.
This yielded \nautorotnumeratormatching light curves. 
While subset $\mathcal{A}$ is more complete than subset $\mathcal{B}$,
this completeness comes at the expense of greater contamination.
% For instance, stars above the slow sequence (possible field
% contaminants) for the most part disappear in subset $\mathcal{B}$

The comparison to the Pleiades and Praesepe is made using the rotation
periods and dereddened colors from Table~5 of
\citet{curtis_rup147_2020}, which used data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.
A few conclusions can be drawn by comparing the \cn\ rotation periods
with those of the Pleiades and Praesepe.  The first is that in both
stages of cleaning, the core and halo of \cn\ each show
gyrochronological ``slow sequences''.  These sequences appear to be
coeval with each other, and at first order appear to overlap
with the Pleiades sequence.

% Second, subset $\mathcal{A}$ is more complete than subset
% $\mathcal{B}$.  This completeness comes at the expense of greater
% contamination.  The slow sequence has a factor of $\sim$twice as many
% stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
% the slow sequence, which could be field contaminants, for the
% most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
% the slow sequence is also more pronounced in the cleaner subset,
% though it seems to not be as empty as in the Pleiades.

%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
Comparing the slow sequence of the Pleiades and \cn\ 
more closely, the top row of Figure~\ref{fig:rot} shows that the two
sequences
overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
The maximum rotation periods seen in \cn, in both subsets
$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
than the $\approx$8.5day limit seen in the Pleiades.  This is
consistent with \cn\ being slightly older than the Pleiades.  Fitting
a model to substantiate this statement \citep[{\it e.g.}, any
of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020} is
an exercise in tautology, since the aforementioned models are empirically fit using
the Pleiades and Praesepe.  In other words, at $(Bp-Rp)_0=1.35$, we
see rotation periods in \cn\ ranging from 7 to 10 days.  The 7 day
periods overlap with the Pleiades; the 10 day periods do not.  Fitting
formulae from the previously cited studies give ages for
a star with $(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}$ of 204 Myr,
316 Myr, and 107 Myr respectively.  Therefore, only the
\citet{spada_competing_2020} model is well-calibrated at these ages
and colors.  For an 8 day rotation period of a star with the same
color, this model quotes a 193 Myr age; at 9 days, 388 Myr.
Given the model uncertainty, we prefer the relative statements that
\cn\ appears slightly gyrochronologically older than the Pleaides, and
much younger than Praesepe.
